Bayesian regression facilitates quantitative modelling of cell metabolism

### 2023-10-31 paper/main.qmd
change condition 2 to dataset 12

To demonstrate how Maud's output can be used to yield useful metabolic insights
we used the results of our case study to explain why the flux of the enzyme
$GNMT$ is higher in condition 1 than in condition 2. GNMT is an irreversible
enzyme that is homotropically activated by its substrate, competitively
inhibited by its product and heterotropically inhibited. The diverse regulation
makes it the ideal test case to elucidate regulatory changes.

@fig-decomposition shows the regulatory description of GNMT, according to the
results our main case study analysis. Each curve shows the marginal posterior
distribution of the ratio of the corresponding regulatory component in condition
1 compared with condition 2. A positive value indicates that the component
was increased in condition 1 relative to condition 2, with 0 indicating no
difference. The probability, according to our model, of the component acting in
each direction is given by the relative area under the curve on each side of the
0 point.

Our model correctly inferred that saturation and allosteric effects were the
main drivers of regulation between the two conditions in this case, with the
curves for each component aligning with the ground truth shown in red.

Importantly, this form of analysis takes into account all modelled sources of
uncertainty, including uncertainty about the true values of the flux in each
condition, and propagates this uncertainty into the final conclusion. Our result
shows that Maud could be used in this realistic case not only to provide a user
with an explanation for an observed difference in fluxes, but also a reasonable
judgement as to the explanation's robustness.

>>>>

To demonstrate how Maud's output can be used to yield useful metabolic
insights we used the results of our case study to explain why the flux of
the enzyme $GNMT$ is higher in dataset 1 than in dataset 12. GNMT is
an irreversible enzyme that is homotropically activated by its substrate,
competitively inhibited by its product and heterotropically inhibited by 5,10-
methylenetetrahydrofolate (mlthf). The diverse regulation makes it the ideal
test case to elucidate regulatory changes.

@fig-decomposition shows the regulatory description of GNMT, according to the
results our main case study analysis. Each curve shows the marginal posterior
distribution of the ratio of the corresponding regulatory component in dataset
1 compared with dataset 12. A positive value indicates that the component
was increased in dataset 1 relative to dataset 12, with 0 indicating no
difference. The probability, according to our model, of the component acting in
each direction is given by the relative area under the curve on each side of the
0 point.

Our model correctly inferred that saturation and allosteric effects were the
main drivers of regulation between the two datasets in this case, with the
curves for each component aligning with the ground truth shown in red.

Importantly, this form of analysis takes into account all modelled sources of
uncertainty, including uncertainty about the true values of the flux in each
dataset, and propagates this uncertainty into the final conclusion. Our result
shows that Maud could be used in this realistic case not only to provide a user
with an explanation for an observed difference in fluxes, but also a reasonable
judgement as to the explanation's robustness.
